Baseline centering problem

### 2018-11-25 paper/tangotd.tex

We assume we can sample transitions $(s,a,r,s')$ according to some
distribution $\rho$ on $s$. In the on-policy case, $\rho$ will be the
stationary distribution. Typically $a$ has to be taken from $\pi$ (or a
distribution close to $\pi$ after resampling/by using only those $a$
sampled from $\pi$; this covers $\eps$-greedy) so that we do get
transitions from
$P$. (Alternatively one could work with pairs $(s,a)$ TODO.) Thus, we
have access to unbiased samples of $\rho P$ and $\rho R$:
\begin{equation}
\rho P=\E[\1_{s}\transp{\1_{s'}}\,],\qquad \rho R=\E[r(s,a,s')\1_{s}]
\end{equation}

If we parameterize $V$ by some parameter $\theta$, we can apply
parametric TD. The expected TD step is
\begin{equation}
\theta\gets \theta -\eta \,\transp{\Phi}\rho (V-PV-R)
\end{equation}
where 
\begin{equation}
\Phi\deq \frac{\partial V}{\partial \theta}
\end{equation}
is the Jacobian matrix of $V$ with respect to its parameter. The linear
case $V=\Phi \theta$ corresponds to constant $\Phi$. The tabular case is
$\Phi=\Id$.

@@

A local minimizer of the $L^2(\rho)$ error between $V$ and
$(\Id-P)^{-1}R$ solves
\begin{equation}
\transp{\Phi} \rho V=\transp{\Phi}\rho (\Id-P)^{-1} R
\end{equation}
which is not quite the same as the above. A local minimizer of the
$L^2(\rho)$ error between $(\Id-P)V$ and $R$ is
\begin{equation}
\transp{\Phi}\transp{(\Id-P)}\rho(\Id-P)V=\transp{\Phi}\transp{(\Id-P)}\rho R
\end{equation}
and in the linear case $V=\Phi\theta$, this is the standard solution to
the linear regression problem $(\Id-P)\Phi\theta\approx R$ in
$L^2(\rho)$.

@@

Only $B=A^{-1}$ and $B=\transp{A}(A\transp{A})^{-1}$ are truly
second-order (namely, solve the problem in one step in the non-stochastic
case). The other algorithms involving an inverse basically have the cost
of second-order algorithms without the full benefits. They are
second-order in feature-space (they orthonormalize $\Phi$) but not in
Markov process space (they don't backpropagate rewards).

>>>>

We assume we can sample transitions $(s,a,r,s')$ according to some
distribution $\rho$ on $s$. In the on-policy case, $\rho$ will be the
stationary distribution. Typically $a$ has to be taken from $\pi$ (or a
distribution close to $\pi$ after resampling/by using only those $a$
sampled from $\pi$; this covers $\eps$-greedy) so that we do get
transitions from
$P$. (Alternatively one could work with pairs $(s,a)$ TODO.) Thus, we
have access to unbiased samples of $\rho P$ and $\rho R$:
\begin{equation}
\rho P=\E[\1_{s}\transp{\1_{s'}}\,],\qquad \rho R=\E[r(s,a,s')\1_{s}]
\end{equation}
and thus also to samples of $\rho(\Id-P)$. Thus, rewriting the equation
as
\begin{equation}
\rho(\Id-P)V=\rho R
\end{equation}
(equivalent if $\rho$ is positive), both sides are samplable, and we can try to solve the above by
stochastic gradient-like methods.

For instance, ff we parameterize $V$ by some parameter $\theta$, we can apply
parametric TD. The expected TD step is
\begin{equation}
\theta\gets \theta -\eta \,\transp{\Phi}\rho (V-PV-R)
\end{equation}
where 
\begin{equation}
\Phi\deq \frac{\partial V}{\partial \theta}
\end{equation}
is the Jacobian matrix of $V$ with respect to its parameter. The linear
case $V=\Phi \theta$ corresponds to constant $\Phi$. The tabular case is
$\Phi=\Id$.

@@

A local minimizer of the $L^2(\rho)$ error between $V$ and
$(\Id-P)^{-1}R$ solves
\begin{equation}
\transp{\Phi} \rho V=\transp{\Phi}\rho (\Id-P)^{-1} R
\end{equation}
which is not quite the same as the above. A local minimizer of the
$L^2(\rho)$ error between $(\Id-P)V$ and $R$ is
\begin{equation}
\transp{\Phi}\transp{(\Id-P)}\rho(\Id-P)V=\transp{\Phi}\transp{(\Id-P)}\rho R
\end{equation}
and in the linear case $V=\Phi\theta$, this is the standard solution to
the linear regression problem $(\Id-P)\Phi\theta\approx R$ in
$L^2(\rho)$. However, the matrix on the left is not samplable.

@@

Only $B=A^{-1}$ and $B=\transp{A}(A\transp{A})^{-1}$ are truly
second-order (namely, solve the problem in one step in the non-stochastic
case).

The other algorithms involving an inverse basically have the cost of
second-order algorithms without the full benefits.  They are second-order
in feature-space (they orthonormalize $\Phi$) but not in Markov process
space (they don't backpropagate rewards).

TODO sampling: if $A=\E \tilde A$, can do $\transp{(\E \tilde A)}(\E
\tilde A)$ or
$\E(\transp{\tilde A}\tilde A)$, which is not equal to $\transp{A}A$ but
is still definite positive (hence stable). This would avoid some
quadruple-sampling problems with $B=\transp{A}(A\transp{A})^{-1}$. But
breaks second-order correctness?

TODO: define second-order with respect to $\Phi$ (second-order
provided $\transp{\Delta}\Delta=\Id$, eg in a Fourier frequency-adjusted
basis), with respect to the Markov chain (second-order provided
$\transp{\Phi}\Phi=\Id$).
